Subset coercions in \Coq

### 2006-05-03 papers/article/main.tex
Bib fixes

\section{Introduction}
There are many means to program in the \Coq{} environment. One
can write programs as in \ML{} and prove properties about them
separately, loosing the possibility of using dependent types in
specifications, or give a rich type expressing them as a goal and use the
proof tactics to solve it, producing a corresponding program by the
\CurryHoward isomorphism but having much less control on its algorithmic
essence. It is however difficult to mix the two methods
(writing code and prove interactively) using a rich specification. We
propose a solution to overcome this difficulty, based on the \ps{}
\cite{Shankar&Owre:WADT99} technique of \PVS{}
\cite{PVS:Language,PVS:Semantics-TR}. This solution also provides
facilities to express properties with a syntaxicaly more mathematical flavor,
bridging the gap between mathematical convention and type theory.

@@

\Russell{}'s typing system (figure \vref{fig:typing-decl-rules}) is based
on the Calculus of Constructions \cite{CoquandHuet,Barras99}, with sigma types but without universes.
This restriction may be removed in future work, but cause no problem for
our main purpose which is programming. The judgement $`G \typed t : T$
means $t$ is a well-typed term of type $T$ in environment $`G$, and $`G
\typed T \subd U$ means $T$ is equivalent to $U$ in environment $`G$. 
Following the presentation as a pure type system \cite{PTS}, 
the set of sorts $\PTSsorts$ is defined as $\setproptype$. As usual, we
let $s, s_i$ for $i `: `N$ range over sorts. The
axioms are $\PTSaxioms = \{ (\Set, \Type), (\Prop, \Type) \}$ and the
set of rules $\PTSrules$ is defined by the following (functional) relation:
\begin{table}[h]
  \vspace{-1.5em}
  \axiomsd
  \vspace{-2em}
  \caption{Definition of $\PTSrules$}
  \label{fig:PTSrules}
\end{table}

@@

The rule \irule{SubConv} integrates $\beta\pi$-conversion in the
judgment. It is well-known \cite{Luo, Pottinger?} that $\beta\pi$ is
confluent and terminating only on well typed terms, so we have to
ensure that our rules, when read from bottom to top, keep
well-typedness. That's the purpose of
the $`G \typed U : s$ premise of the \irule{SubTrans} rule. This rule
ensures that our judgement constructs an \emph{equivalence} and has
proper modularity. The next two rules do context closure for dependent
products and sums. It is remarkable that we use contravariance
\cite{journals/toplas/Castagna95} for
domains in the \irule{SubProd} rule, not restricting to invariance as in
\PVS. It is accessory here, as we could have used \emph{co}variance and
still get the same judgements because we have symetry. However, it will
become important when we create coercions (see
\vref{fig:subtyping-impl-rules}).

@@

\subsubsection{Properties}
\begin{paragraph}{Assumptions}
  We have proved basic metatheoretical results for this system in \Coq{}
  \cite{CCP}. We have assumed that it is strongly normalizing (SN) and has the
  following subject reduction (SR) property: if $`G \typed t : T$ and $t
  "->"_{\beta\pi} t'$ then $`G \typed t' : T$. Strong normalization
  could be proved using the generic proof method developed in
  \cite{Geuvers} but we leave it for future work. Gang Chen
  \cite{GangChen} has studied various type systems from the
  $\lambda$-cube extended by subtyping or coercive subtyping \cite{Luo},
  including the Calculus of Constructions, and proved such results as
  SN and SR for them. It requires a very careful analysis of the system 
  to avoid cycles in the proof due to the presence of dependent types and
  conversion ; hence we also leave it for future work.
\end{paragraph}

@@

The next thing to do is to construct a decidable judgement for
equivalence given two well-sorted types. We denote it by $`G \typea T
\suba U$ (figure \ref{fig:subta-rules}).
In our case, we have to check that conversion is decidable, which is the
case. We also need to eliminate the transitivity rule which is not
syntax-directed. That is the purpose of the following theorem:
\begin{theorem}[Elimination of transitivity]
  If $`G \typea T \suba U$ and $`G \typea U \suba V$, then $`G \typea T
  \suba V$.
\end{theorem}
\begin{proof}
  By lexicographic induction on the pair of derivations.
  \TODO{example}
\end{proof}

@@

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% LaTeX-command: "TEXINPUTS=\"..:../style:$TEXINPUTS\" latex"
%%% End:

>>>>

\section{Introduction}
There are many means to program in the \Coq{} environment. One
can write programs as in \ML{} and prove properties about them
separately, loosing the possibility of using dependent types in
specifications, or give a rich type expressing them as a goal and use the
proof tactics to solve it, producing a corresponding program by the
\CurryHoward isomorphism but having much less control on its algorithmic
essence. It is however difficult to mix the two methods
(writing code and prove interactively) using a rich specification. We
propose a solution to overcome this difficulty, based on the \ps{}
\cite{Shankar&Owre:WADT99} technique of \PVS{}
\cite{PVS:Language,PVS-Semantics:TR}. This solution also provides
facilities to express properties with a syntaxicaly more mathematical flavor,
bridging the gap between mathematical convention and type theory.

@@

\Russell{}'s typing system (figure \vref{fig:typing-decl-rules}) is based
on the Calculus of Constructions \cite{coquandhuet88,Barras99}, with sigma types but without universes.
This restriction may be removed in future work, but cause no problem for
our main purpose which is programming. The judgement $`G \typed t : T$
means $t$ is a well-typed term of type $T$ in environment $`G$, and $`G
\typed T \subd U$ means $T$ is equivalent to $U$ in environment $`G$. 
Following the presentation as a pure type system \cite{barendregt91}, 
the set of sorts $\PTSsorts$ is defined as $\setproptype$. As usual, we
let $s, s_i$ for $i `: `N$ range over sorts. The
axioms are $\PTSaxioms = \{ (\Set, \Type), (\Prop, \Type) \}$ and the
set of rules $\PTSrules$ is defined by the following (functional) relation:
\begin{table}[h]
  \vspace{-1.5em}
  \axiomsd
  \vspace{-2em}
  \caption{Definition of $\PTSrules$}
  \label{fig:PTSrules}
\end{table}

@@

The rule \irule{SubConv} integrates $\beta\pi$-conversion in the
judgment. It is well-known that $\beta\pi$ is
terminating only on well typed terms, so we have to
ensure that our rules, when read from bottom to top, keep
well-typedness. That's the purpose of
the $`G \typed U : s$ premise of the \irule{SubTrans} rule. This rule
ensures that our judgement constructs an \emph{equivalence} and has
proper modularity. The next two rules do context closure for dependent
products and sums. It is remarkable that we use contravariance
\cite{journals/toplas/Castagna95} for
domains in the \irule{SubProd} rule, not restricting to invariance as in
\PVS. It is accessory here, as we could have used \emph{co}variance and
still get the same judgements because we have symetry. However, it will
become important when we create coercions (see
\vref{fig:subtyping-impl-rules}).

@@

\subsubsection{Properties}
\begin{paragraph}{Assumptions}
  We have proved basic metatheoretical results for this system in \Coq{}
  \cite{CCP}. We have assumed that it is strongly normalizing (SN) and has the
  following subject reduction (SR) property: if $`G \typed t : T$ and $t
  "->"_{\beta\pi} t'$ then $`G \typed t' : T$. Strong normalization
  could be proved using the generic proof method developed in
  \cite{geuvers95} but we leave it for future work. Gang Chen
  \cite{ChenPhD} has studied various type systems from the
  $\lambda$-cube extended by subtyping or coercive subtyping \cite{DBLP:conf/csl/Luo96},
  including the Calculus of Constructions, and proved such results as
  SN and SR for them. It requires a very careful analysis of the system 
  to avoid cycles in the proof due to the presence of dependent types and
  conversion ; hence we also leave it for future work.
\end{paragraph}

@@

The next thing to do is to construct a decidable judgement for
equivalence given two well-sorted types. We denote it by $`G \typea T
\suba U$ (figure \ref{fig:subtyping-algo-rules}).
In our case, we have to check that conversion is decidable, which is the
case. We also need to eliminate the transitivity rule which is not
syntax-directed. That is the purpose of the following theorem:
\begin{theorem}[Elimination of transitivity]
  If $`G \typea T \suba U$ and $`G \typea U \suba V$, then $`G \typea T
  \suba V$.
\end{theorem}
\begin{proof}
  By lexicographic induction on the pair of derivations.
  \TODO{example}
\end{proof}

@@

\bibliography{bib-joehurd,barras,pvs-bib,bcp,Luo,biblio-subset,%
cparent,demons,demons2,demons3}
\bibliographystyle{plain-url}

\renewcommand{\thefootnote}{}
\footnotetext{Ce rapport a été préparé sous \LaTeX~avec la fonte 
  \texttt{Concrete Math}}
\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% LaTeX-command: "BIBINPUTS=\"${HOME}/research/demons-biblio:../bib:$BIBINPUTS\" TEXINPUTS=\"..:../style:$TEXINPUTS\" latex"
%%% End:
